---
title: "Nov-1-2024-practice.qmd"
format: html
editor: visual
---

## Introduction

At this point, you have learned to perform and interpret three different statistical tests: Chi-square test of association (contingency analysis), two-sample t-test and linear regression.

Now it is time for you to practice what you've been learning. Please add the data set called "A_morio_2014_plants.csv" to your Biostats-In-Class RStudio project as you will need it for this exercise. The data are associated with [this paper in the journal Evolution](https://academic.oup.com/evolut/article/70/3/716/6852093?login=true) and come from [this repository on Dryad](https://datadryad.org/stash/dataset/doi:10.5061/dryad.nf663). The authors of the study were interested in the degree to which pollinators apply selective pressure on a variety of flower traits in a species of orchid plant. In the experiment, some of the orchids were allowed to experience natural pollination and were considered the control group, and another group of orchids was pollinated by hand by the researchers (the "hand-pollinated" treatment group).

### Question 1

The researchers wished to [know whether spur length]{.underline} (the "spur" is a part of the flower that sticks out) of flowers [differed between control and hand-pollinated plants]{.underline}.

Your task is to determine the appropriate statistical test you should use to answer this question from the data and then proceed with the statistical analysis, using the "plot your data (and guess the answer!) -\> build your model -\> check your assumptions -\> interpret your model" workflow (for now, we will leave out the last step which is to make a better plot at the end).

I am expecting that you will document your process clearly along the way.

Set up

```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(ggfortify)
```

Read in data

```{r}
orchids <- read.csv(here("Data", "A_morio_2014_plants.csv"))
```

Change data types

```{r}
orchids$treatment <- as.factor(orchids$treatment)
orchids$plant_ht_class <- as.factor(orchids$plant_ht_class)
```

**Stat test to use is a two sample t-test**

Predictor is treatment, response is spur length

First plot the data to look at relationships

```{r}
# Use a histogram and facet wrap for the two treatments

ggplot(orchids, aes(x= spur_length_mm))+
  geom_histogram()+
  facet_wrap(~treatment, ncol=1)

#Seems like there will be a difference, but not huge because there is some overlap, but control extends more
```

Now look at means of each group

```{r}
mean.spur <- orchids |>
  group_by(treatment) |>
  summarise(mean.spur = mean(spur_length_mm, na.rm=T), stdev.spur = sd(spur_length_mm, na.rm=T))
```

Means seem very similar, so may not be a difference

Run the t-test

```{r}
t.test(spur_length_mm ~ treatment, data=orchids)
```

p-value is 0.03666 which is significant depending on what level you consider significant

t = 2.0987 and df = 309.83

There is a signficant difference in the mean spur length between the control treatment and the hand-pollinated treatment, with a greater spur length in the control group.

### Question 2

The researchers wished to know if there was an [association between plant height class]{.underline} and [treatment group]{.underline} and the [number of flowers]{.underline} produced by orchid plants.

Your task is to determine the appropriate statistical test you should use to answer this question from the data and then proceed with the statistical analysis, using the "plot your data (and guess the answer!) -\> build your model -\> check your assumptions -\> interpret your model" workflow (for now, we will leave out the last step which is to make a better plot at the end).

I am expecting that you will document your process clearly along the way.

Stat test - regression

Plot with a scatterplot

```{r}
ggplot(orchids, aes(x = ))
```

### Question 3

The researchers wished to know if there is an association between corolla area and spur length in hand-pollinated flowers.

Your task is to determine the appropriate statistical test you should use to answer this question from the data and then proceed with the statistical analysis, using the "plot your data (and guess the answer!) -\> build your model -\> check your assumptions -\> interpret your model" workflow (for now, we will leave out the last step which is to make a better plot at the end).

I am expecting that you will document your process clearly along the way.
