---
title: "Practice with t-tests"
output: html_document
date: "2022-10-27"
---

## Reviewing the t-test

Let's conduct a 2-sample t-test to compare body fat percentage between men and women.

The data set presents the body fat percentage for a sample of men and women who all work out at the same gym. In general, men are leaner than women, with a body fat percentage centering around 15% whereas for women it tends to center around 20%. Let’s see if the same is true of these data.

## Instructions

Use the information you learned from reading chapter 5 and doing homework 21 to conduct a t-test comparing mean body fat percentage between the male and female groups. But don’t just rush to the test – follow the recommended analysis workflow:

1.  Plot the data and look at expected relationships
2.  Calculate the mean and standard deviation for the data and examine the resulting output.
3.  Discuss with your partner:
    -   Review in the book: What are the assumptions about the data appropriate for a t-test? When you examine the plots and summary data, do they appear to meet those assumptions?
        -   *Data is normally distributed*
        -   *Variances are equal*
        -   *Sample sizes are small*
    -   Based on examination of the plots and summary data, do you expect that there is a statistically significant difference in mean body fat percentage between males and females in this sample?
        -   *Yes, means are pretty different, and histogram doesn't overlap too much*
4.  Conduct a t-test and then interpret the results. What do they tell you about the null vs. alternative hypotheses?

```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

```{r}
body.fat <- read.csv(here("Data", "body_fat_percentage.csv"))
```

```{r}
body.fat$gender <- as.factor(body.fat$gender)
glimpse(body.fat)
```

```{r}
ggplot(body.fat, aes(x = body_fat_percent))+
  geom_histogram(binwidth=10)+
  facet_wrap(~ gender, ncol=1)+
  theme_bw()
```

```{r}
mean.body.fat <- body.fat |>
  group_by(gender) |>
  summarise(mean.body.fat = mean(body_fat_percent), stdev.body.fat = sd(body_fat_percent))
```

```{r}
t.test(body_fat_percent ~ gender, data = body.fat)
```

*Our results show that there is a significant difference between the means in body fat percentage of the female and male groups, so we can reject the null hypothesis.*
