---
title: "Nov 4 Class"
format: html
editor: visual
---

## ANOVA tree data

```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(ggfortify)
```

```{r}
trees <- read.csv(here("Data", "Forest_Tree_DBH_F21.csv"))
```

```{r}
glimpse(trees)
trees$Forest <- as.factor(trees$Forest)
```

Part 1 - Is there a difference in the mean DBH between the forests?

Part 2 - If there is a difference, for which forests does the difference exist?

## Part 1

Make a plot

```{r}
ggplot(trees, aes(x=Forest, y=DBH_cm))+
  geom_boxplot()+
  theme_bw()

#geom_jitter -> erika used and makes points not on top of each other
```

Seems like the Degrasse has a higher mean than the others.

```{r}
model_trees <- lm(DBH_cm ~ Forest, data=trees)
```

```{r}
autoplot(model_trees, smooth.colour = NA)
```

```{r}
anova(model_trees)
```

Mean DBH is not the same in the forests (significant).

```{r}
summary(model_trees)
```

The Intercept group is Degrasse because it's the first alphabetically so shows diff b/w that and the others

## Part 2

Install and activate the multcomp package

```{r}
library(multcomp)
```

Conduct a Tukey's HSD (honestly significant differences) test

-   Only run a Tukey's test if you get a significant p-value with the ANOVA on its own

<!-- -->

-   Compares all the forests with all the other forests

-   Automatically only gives significance once it's adjusted the p values

```{r}
phc1 <- glht(model_trees, linfct = mcp(Forest = "Tukey"))
summary(phc1)


#glht(the model, linear factor = multiple comparison(give factor variable = "Tukey"))
#mcp means multiple comparisons

#use summary(__) to get results
```

This doesn't compare against an intercept

== 0 is the null (no difference)

Can act for a 'compact letter display' of the results – tells differences based on letters. Any group that shares the same letter are not different from each other

```{r}
cld(phc1)
```

Degrasse and South Hammond don't share the same letter
