---
title: "Nov 20 Class - Casting and joining"
format: html
editor: visual
---

# Casting b/w long –\> wide and joining dataframes

Set up

```{r}
rm(list=ls())
library(tidyverse)
library(here)

```

read in data 1

```{r}
cams <- read.csv(here("Data", "site_locations.csv"))
```

read in data 2

```{r}
species <- read.csv(here("Data", "species_per_site.csv"))
```

## Casting b/w long and wide

Might want to make a column for all the dates - because might want to make a plot with just one date on the x or y axis (can only do with a column)

Use `pivot_wider()` to do this

```{r}
sp_wide <- species |> pivot_wider(names_from = Date, values_from = Num_species)
```

End up with NAs because signifies days that camera wasn't out at that site

Can go other direction and cast with `pivot_longer()`

```{r}
sp_long <- sp_wide |>
  pivot_longer(cols = c(`1-Jul-24`:`15-Jul-24`), names_to = "Date", values_to = "Num_species")
```

\` says to R to think of the value as text

the : says to go from here to here

End up with more rows because it incorporated the NAs - which is fine

## Joining Dataframes

How to combine two dataframe that are different lengths:

Want to get the lat and long data from cams and add it to species:

They both share the site variable, so can use that to match

Use `join` function to do it

```{r}
cams_species <- full_join(species, cams) 
#First data frame is the one that will be added to - in this case, cams is being added to species

# by = ___ would tell what column you want to match, don't need if only one thing in common

```

The help panel has more options

Keeps all 81 rows of data because it's a full join - need to be thoughtful about which join you use.

```{r}
site_species <- full_join(cams, species)
# In this case, species is being added to cams
```
