---
title: "Oct 3 2024 Class"
format: html
editor: visual
---

## Set up

```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

Now read in data:

```{r}
drugs <- read.csv(here("Data", "drug_resistance.csv"))
```

```{r}
str(drugs)
glimpse(drugs)
```

turn treatment variable into a factor

```{r}
drugs$Treatment <- as.factor(drugs$Treatment)

levels(drugs$Treatment)
```

There are 9 levels b/c no data validation used - should be 3 levels.

Now fix the treatment variable

Find the rows that have any version of "control"

Use 'which()' function to do so.

```{r}
fix <- which(drugs$Treatment %in% c("cantrol", "control", "Control")) 
#look in drugs treatment column and return row numbers where value in column is included in this list of options (and call that list fix)
#identified the rows that need repair
```

Now fix

```{r}
drugs$Treatment[fix] <- "control"

levels(drugs$Treatment) #gives the same thing so must ask it to relist the levels...

drugs$Treatment <-factor(drugs$Treatment) #this just tells R to re-list the levels of my factor now that they have changed

levels(drugs$Treatment) #now shows just one version of control variable
```

Now fix the rest of the column Treatment to have only 3 levels

Fix drug 2

```{r}
fix2 <- which(drugs$Treatment %in% c("drug x 2", "Drugx2", "drug2"))

drugs$Treatment[fix2] <- "drug x 2"

drugs$Treatment <-factor(drugs$Treatment)
levels(drugs$Treatment)
```

Now fix drug 3

```{r}
fix3 <- which(drugs$Treatment %in% c("drug x 3", "Drug x 3", "drug3"))

drugs$Treatment[fix3] <- "drug x 3"

drugs$Treatment <-factor(drugs$Treatment)
levels(drugs$Treatment)
```

## Now fix tech.name column

First make a factor and look at levels

```{r}
drugs$tech.name <- as.factor(drugs$tech.name)

levels(drugs$tech.name)
```

Now fix to have right number of levels (4)

```{r}
FixBecca <- which(drugs$tech.name %in% c("becca", "Becca"))

drugs$tech.name[FixBecca] <- "Becca"

drugs$tech.name <-factor(drugs$tech.name)
levels(drugs$tech.name)
```

```{r}
FixBrian <- which(drugs$tech.name %in% c("brain", "brian", "Brian"))

drugs$tech.name[FixBrian] <- "Brian"

drugs$tech.name <-factor(drugs$tech.name)
levels(drugs$tech.name)
```

```{r}
FixJose <- which(drugs$tech.name %in% c("jose"))

drugs$tech.name[FixJose] <- "Jose"
```
