---
title: "rabies.qmd"
format: html
editor: visual
---

## Rabies project

You are receiving a dataset from the New York State Department of Health that shows test results for specimens submitted to be tested for rabies.

## In Class Exercise 1
### Set things up and read in the data

Add the code to set up the workspace

```{r}

```

Now read in the data and name it as an object called `rabies`.

```{r}

```

Reduce the dataset to just the first 7 columns

```{r}

```

### Examine the data

1.  Get a summary for each variable.

```{r}

```

2.  Repair column names

```{r}

```

3.  Adjust data types

```{r}

```

4.  Now plot some data

```{r}

```
## In Class Exercise 2
### Look at some relationships
5. Create a percent positive column
```{r}

```

6. Using the rabies dataset, create a new object called "animal summary" in which you determine, by Year and Animal Classification, the total number of samples, total positive samples, and the mean percent positive samples.

```{r}
animal_summary <- rabies %>% group_by(Received_Year, Animal_Classification) %>% summarise(
  Number_Samples = sum(Number_Samples),
  Positive_Samples = sum(Positive_Samples),
  mean_percent_pos = mean(percent_pos)
  
)


```
Next, make a boxplot of those data with the following features:
- The x-axis should be the animal classificaton
- The y-axis should be the mean percent positive samples
- Your boxplot should include the individual points laid over the boxplots. Make the points gray.
- Rename the x-axis label as "Animal type" and the y-axis label as "Mean percent positive samples"


```{r}

```

